<!--- Below are global settings for knitr. You can override any of them by adding the changes to individual chunks --->

```{r global_options, include=FALSE}
knitr::opts_chunk$set(error = TRUE,         # Keep compiling upon error
                      collapse = FALSE,     # collapse by default
                      echo = TRUE,          # echo code by default
                      comment = "#>",       # change comment character
                      fig.width = 5,        # set figure width
                      fig.align = "center", # set figure position
                      out.width = "49%",    # set width of displayed images
                      warning = FALSE,      # do not show R warnings
                      message = FALSE)      # do not show R messages
```

<!--- Solution Region --->
```{css, echo=FALSE}
#solution {
  background-color: #8FBC8F;
  border-style: solid;
  border-color: blue;
  padding: .5em;
  margin: 20px
}
```


**SYS 6018 | Fall 2020 | University of Virginia **

*******************************************

<!--- Load Required R packages here --->
```{r, include=FALSE}
library(tidyverse)
digits <- function(x, k=2) format(round(x, k), nsmall=k)
data.dir = 'https://mdporter.github.io/SYS6018/data/'
```

### Problem 0.1: Who are you?

a. What is your name? 

<div id="solution">
Congxin (David) Xu
</div>

b. Where are you from (e.g., hometown)? 

<div id="solution">
Hefei, Anhui, China
</div>

c. What do you hope to do after graduation (be specific)?  

<div id="solution">
Work as a Lead/Senior Data Scientist
</div>

d. List 3 interesting things about yourself. 

<div id="solution">
1. I love playing board games
2. I love playing sports with a racquet, such as badminton, tennis, racquetball, squash, etc.  
3. I have 3-year-old American short hair cat, Maca. 
</div>


----------

### Problem 0.2: About the course

a. What are you most excited to learn in this course?

<div id="solution">
Study different data mining techniques and understand how those techniques can be applied to the machine learning forecasting models. 
</div>


b. What pre-req topics (e.g., math, coding, probability, statistics) are your weakest? 

<div id="solution">
I believe math may be my weakest pre-req topic. 
</div>


c. What pre-req topics are your strongest? 

<div id="solution">
I think I am strong at coding, probability and statistics.
</div>

d. Would you rather see the mathematical details or coding details of a method/algorithm? 

<div id="solution">
YES! Those are exactly what I am looking for in our MSDS program.
</div>

----------

### Problem 0.3: Math Notation

a. What are the equations for the least squares coefficients in linear regression (in matrix notation)? Use $X$ for the design/predictor matrix, and $Y$ the vector of responses. 

<div id="solution">
Using the matrix notation, the equation for the least squares coefficients in linear regression is 
$$ Y = X \pmb{\beta} + \pmb{\varepsilon} $$
where 
$$ Y = \begin{bmatrix} y_1 \\ y_2 \\ \vdots \\ y_n 
       \end{bmatrix}, \enspace
X = \begin{bmatrix} 1 & x_{11} & x_{12} & \ldots & x_{1k} \\ 
                    1 & x_{21} & x_{22} & \ldots & x_{2k} \\ 
                    \vdots & \vdots & \vdots &   & \vdots \\ 
                    1 & x_{n1} & x_{n2} & \ldots & x_{nk} \\ 
    \end{bmatrix} , \enspace 
\pmb{\beta} = \begin{bmatrix} \beta_0 \\ \beta_1 \\ \vdots \\ \beta_k 
              \end{bmatrix}, \enspace
\pmb{\varepsilon} = \begin{bmatrix} \varepsilon_1 \\ \varepsilon_2 \\ \vdots \\ \varepsilon_n                       \end{bmatrix} $$
</div>


b. Let $x_1, x_2, \ldots, x_n$ be a sample of length of time that a customer is on the phone with a call center help line. We feel comfortable modeling the data as coming from an *exponential distribution*. What is the MLE (Maximum Likelihood Estimate) of the parameter?  Show your steps.

<div id="solution">
The probability density function of the exponential distribution is defined as

$$
f(x;\lambda) = \begin{cases}
\lambda e^{-\lambda x} & \text{if } x \geq 0 \\
0 & \text{if } x<0
\end{cases}
$$

Its likelihood function is

$$
\mathcal{L}(\lambda, x_1, \dots, x_n) = \prod_{i = 1}^n f(x_i, \lambda) = \prod_{i = 1}^n \lambda e^{-\lambda x} = \lambda^n e^{-\lambda\sum_{i=1}^n x_i}
$$

To calculate the maximum likelihood estimator I solved the equation

$$
\frac{d\ln\left(\mathcal{L}(\lambda,x_1,\dots,x_n)\right)}{d\lambda}\overset{!}{=}0
$$

for $\lambda$.

$$
\begin{align}
\frac{d\ln\left(\mathcal{L}(\lambda,x_1,\dots,x_n)\right)}{d\lambda}
&= \frac{d\ln\left(\lambda^ne^{-\lambda\sum_{i=1}^nx_i}\right)}{d\lambda} \\
&= \frac{d\ln\left(n\ln(\lambda)-\lambda\sum_{i=1}^n x_i\right)}{d\lambda} \\
&= \frac{n}{\lambda}-\sum_{i=1}^n x_i
\end{align}
$$

Finally we get $$\lambda = \frac{n}{\sum\limits_{i=1}^n x_i}$$


*Resource*: https://math.stackexchange.com/questions/101481/calculating-maximum-likelihood-estimation-of-the-exponential-distribution-and-pr

</div>


### Problem 0.4: Coding

a. Simulate 100 observations from the following model:

- $X \sim N(1, 1)$
- $Y \sim N(1 + 2X, 2)$
- $Z = \begin{cases} 1 &\quad Y<0 \\ 2 &\quad Y \ge 0 \end{cases}$

<div id="solution">


```{r}
set.seed(666)
x <- rnorm(n = 100, mean = 1, sd = 1)
y <- rnorm(n = 100, mean = 1 + 2*x, sd = 2)
z <- ifelse(y < 0, 1, 2)
```

</div>

b. Make a scatter plot of the data. Put $X$ on the x-axis and $Y$ on the y-axis. Color the points according to $Z$. 

<div id="solution">
```{r}
library(ggplot2)
df <- data.frame(x, y, z)
ggplot(data = df) + geom_point(aes(x = x, y = y, color = z))
```
</div>